A ``Dream'' Interface Design

### 2014-11-25 dream-design/dream-design.tex
more edits

\begin{abstract}
This investigation presents a ``dream'' user interface design for touch screen mobile devices. This design aims to tackle the usability issue of insufficient feedback as well as provide an interface that is user-friendly for individuals with visual impairments and individuals without visual impairments alike. Of the five prevalent usability metrics within the field of Interaction Design, the three metrics of learnability, efficiency, and errors were particularly significant in the creation and development of this interface design \cite{Nielsen:1993:UE:529793}.
\end{abstract}

@@

Many of these mobile devices today utilize touch screens instead of physical keyboards or buttons. This allows for a larger display without increasing the size of the device. Furthermore, touchscreen keyboards and buttons have the ability to change their layout based on user input and disappear when not needed. To compensate for the lack of tactile feedback provided by physical keys and buttons, touchscreen devices often include aural feedback in the form of audible clicks from a speaker. Haptic feedback in the form of device vibrations is often included as well.

Even with these alternate forms of feedback, the literature suggests that insufficient feedback is still a major usability issue with touchscreen mobile devices. \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}. A study by Hussain Tinwala and Scott MacKenzie submits that the lack of physical keys requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}. The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users. A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the interface or the screen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.

Thus, the aim of this investigation is to propose a ``dream'' interface design that addresses this usability issue.

@@

My ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology \cite{Tactus}. Tactus Technology, a company based in Fremont, California, creates real physical buttons that dynamically appear and disappear into a flat touch screen (see Figure~\ref{tactus1}). Small fluid channels are routed throughout the Tactile Layer and enable fluid to expand the top polymer layer to create the physical buttons \cite{Tactus}.

Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition as well as several awards \cite{CNN, I-Zone, PCMag, Wired}. Reviewers have articulated how the technology seems to be ``downright magical'' \cite{CNN}. Though the look and feel of Tatctus technology seems totally futuristic, the technology is already beginning to appear in consumer devices. In 2013, Touch Revolution, the largest-volume glass projected capacitive multi-touch screen manufacturer in the world, announced a partnership with Tactus Technology \cite{TactusAvailability}. The technologies from the two companies are being combined and manufactured into consumer devices today \cite{TactusAvailability}. Not only is this technology becoming more and more available, but it is becoming more customizable as well. Companies will soon be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote \cite{CNN}.

@@

\section{Top-Level Design}
The objective of this ``dream'' interface design is to make mobile touchscreen devices more usable for individuals with visual impairments and individuals without visual impairments alike. The three main features of the design used to attempt to accomplish this goal are the optional physical grid, the ``intelligent'' tactile buttons, and the aural feedback.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=2in]{wireframe-grid.png} 
\caption{A simple wireframe demonstrating the concept of the grid layout. Each line on the screen of the figure would actually be a physical line that appears on the device's screen with the use of Tactus's microfluidic technology.}
\label{wireframe-grid}
\end{figure}

\subsection{``Intelligent'' Tactile Buttons}
The design also includes tactile buttons that would rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}). Like with the optional physical grid, this aspect of the interface design would also make use of Tactus's microfluidic technology.

Ideally, only buttons applicable to the input of the user or current state of the device would appear. For example, in an instance where the user is provided with a binary yes/no choice, only two tactile buttons should appear---one for yes and one for no. Similarly, if all possible buttons were applicable for a given situation, then all possible buttons should appear.

@@

\subsection{Aural Feedback}
The design of this interface was created in such a way that the tactile aspects are intended to operate in conjunction with other technologies that provide aural feedback. Since the grid and tactile buttons can only help the user know \textit{where} an application is, this aural feedback can inform the user of exactly \textit{what} the application is.

This portion of the interface design should not be too difficult to implement as there already exists technology that provides this kind of aural feedback. For example, the Apple operating system, iOS, has the VoiceOver screen reader that provides a spoken description of everything happening on the screen \cite{VoiceOver}. The Android OS has a similar service called TalkBack, which adds spoken and audible feedback to the device \cite{TalkBack}.

\section{Usage Scenarios}
There exist several usage scenarios for this interface design. Sighted users employ vision to tell the user what and where to touch, but this interface lacks accessibility for IBSVI because it lacks physical landmarks and tangible buttons. A typical solution is to provide voice over to provide this information, but there is nothing to guide IBSVI to that location in the first place, and there are no landmarks other than dead-reckoning and pure spatial memory to support navigation for the IBSVI. For example, the Apple operating system, iOS, has VoiceOver function, and the Android OS has a Talkbalk function. iOS VoiceOver reads aloud the icon, the user should double tap on that location without the benefit of vision to ensure that the finger does not stray from the touch point.

The problem with such accessibility modes is that the user can develop a spatial mental model for the interface or screed only through dead reckoning to find VoiceOver locations, in the absence of any landmark other than the boundary of the device. The optional tactile grid as well as the three-dimensional buttons that appear with this design solves this problem. The additional tactile feedback and physical landmarks can guide IBSVI to interact efficiently and effectively with the touch device and its voice over functionality.

Another use case for this interface pertains specifically to the ``intelligent'' tactile buttons. Users often may get overwhelmed by the number of options that are always available to them. It also may be frustrating for users to try to perform a task and not be able to. To examine a specific example, let's say that the user cannot take any more photos on their phone since they are out of memory. In this case, the interface would not even raise the button used to capture photos. This is a good indicator to the user that they need to perform some kind of an action before proceeding.

>>>>

\begin{abstract}
This investigation presents a ``dream'' user interface design for touchscreen mobile devices. This design aims to tackle the touchscreen usability issue of insufficient feedback as well as provide an interface that is user-friendly for individuals with visual impairments and individuals without visual impairments alike. Of the five prevalent usability metrics within the field of Interaction Design, the three metrics of learnability, efficiency, and errors were particularly significant in the creation and development of this interface design \cite{Nielsen:1993:UE:529793}.
\end{abstract}

@@

Many of these mobile devices today utilize touchscreens instead of physical keyboards or buttons. This allows for a larger display without increasing the size of the device. Furthermore, touchscreen keyboards and buttons have the ability to change their layout based on user input and disappear when not needed. To compensate for the lack of tactile feedback provided by physical keys and buttons, touchscreen devices often include aural feedback in the form of audible clicks from a speaker. Haptic feedback in the form of device vibrations is often included as well.

Even with these alternate forms of feedback, the literature suggests that insufficient feedback is still a major usability issue with touchscreen devices. \cite{Tinwala:2010:ETE:18	68914.1868972, Kane:2011:UGB:1978942.1979001, Hardy:2008:TIT:1409240.1409267, El-Glaly:2013:TTF:2460625.2460665, Buxton:1986:HID:22339.22386}. A study by Hussain Tinwala and Scott MacKenzie submits that the lack of physical keys requires heightened visual attention from the user, which diverts the user's concentration from the thoughts being expressed \cite{Tinwala:2010:ETE:18 68914.1868972}. The lack of physical keys not only diverts the attention of some users, but it also makes the device  almost entirely unusable for other users. A study from Virginia Polytechnic Institute and State University demonstrates how touchscreen mobile devices do not provide sufficient feedback for Individuals with Blindness or Severe Visual Impairment (IBSVI) since these users are only able to develop a spatial mental model for the touchscreen through ``dead reckoning'' \cite{El-Glaly:2013:TTF:2460625.2460665}.

Thus, the aim of this investigation is to propose a ``dream'' interface design that addresses the usability issue of insufficient feedback.

@@

My ``dream'' interface design incorporates functionality of a product that is currently being developed by Tactus Technology \cite{Tactus}. Tactus Technology, a company based in Fremont, California, creates technology that can reshape the surface of touchscreens on demand. The technology allows real physical buttons to dynamically appear and disappear into a flat touchscreen (see Figure~\ref{tactus1}). Small fluid channels are routed throughout the Tactile Layer and enable fluid to expand the top polymer layer to create the physical buttons \cite{Tactus}.

Although this technology from Tactus is extremely bleeding-edge, the prototypes have already received a fair amount of recognition \cite{CNN, I-Zone, PCMag, Wired}. Reviewers have articulated how the technology seems to be ``downright magical'' \cite{CNN}. Though the look and feel of Tatctus technology seems totally futuristic, the technology is already beginning to appear in consumer devices. In 2013, Touch Revolution, the largest-volume glass projected capacitive multi-touch screen manufacturer in the world, announced a partnership with Tactus Technology \cite{TactusAvailability}. The technologies from the two companies are being combined and manufactured into consumer devices today \cite{TactusAvailability}. Not only is this technology becoming more and more available, but it is becoming more customizable as well. Companies will soon be able to customize the panel for different types of buttons, say for example, the buttons on a TV remote \cite{CNN}.

@@

\section{Top-Level Design}
The objective of this ``dream'' interface design is to make mobile touchscreen devices more usable for individuals with visual impairments and individuals without visual impairments alike. The three main features of this design are the optional physical grid, the ``intelligent'' tactile buttons, and the aural feedback.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=2in]{wireframe-grid.png} 
\caption{A simple wireframe demonstrating the concept of the grid layout. Each line in the figure would actually be a physical line that appears on the device's screen with the use of Tactus's microfluidic technology.}
\label{wireframe-grid}
\end{figure}

\subsection{``Intelligent'' Tactile Buttons}
The interface design also includes tactile buttons that use Tactus's microfluidic technology to rise up and recede in an ``intelligent'' manner (see Figure~\ref{buttons}). Ideally, only buttons applicable to the input of the user or current state of the device would appear at any given time. For example, in an instance where the user is provided with a binary yes/no choice, only two tactile buttons should appear---one for yes and one for no. Similarly, if all possible buttons were applicable for a given situation, then all possible buttons should appear.

@@

\subsection{Aural Feedback}
\label{aural-feedback}
The design of this interface was created in such a way that the tactile aspects are intended to operate in conjunction with other technologies that provide aural feedback. Since the grid and tactile buttons can only help the user know \textit{where} an application is, this aural feedback can inform the user of exactly \textit{what} an application is.

This type of feedback should not be too difficult to implement as there already exists technology that provides similar aural feedback. For example, the Apple operating system, iOS, has the VoiceOver screen reader that provides a spoken description of everything happening on the screen \cite{VoiceOver}. The Android OS has a similar service called TalkBack, which adds spoken and audible feedback to the device \cite{TalkBack}.

\section{Usage Scenarios}
There exist several usage scenarios for this interface design. One such scenario is when an Individual with Blindness of Servere Visual Impairment (IBSVI) attempts to locate an application on a touchscreen mobile device. Sighted users employ vision to figure out what and where to touch. Since IBSVI cannot do this, typical touchscreen devices lack accessibility for IBSVI. As previously touched on in Section~\ref{aural-feedback}, a typical solution to this problem is to use voice over technology to provide information about what is on the screen at any given time. However, there is nothing to guide IBSVI to the location of a specific application in the first place. Since there are no landmarks to support navigation for the IBSVI, the user must use dead-reckoning and trial-and-error methods in order to find applications specified by the voice over technology.

The design of this interface addresses this problem as the tactile grid and the three-dimensional buttons provide additional landmarks for IBSVI. These landmarks aim to allow IBSVI interact more efficiently and effectively with the touch device and its voice over functionality. Ideally, the voice over functionality would be able to tell the user the exact cell in the grid in which the desired application is located. The user could then find the application much more easily.

Another use case for this interface is when some functionality of an application is disabled. Say, for example, that a user cannot take any more photos on their mobile device because they are out of memory. Some ``intelligent'' tactile picture-taking button could recede when this happens to communicate to the user that photo taking functionality is disabled. More generally, the``intelligent'' tactile buttons can be useful in indicating to the user what choices they have at any given time.
